### 2024-08-28 2_developmental_analysis.qmd

---
title: "Trial-level developmental analysis"
format:
  html:
    toc: true
    toc-location: left
---

```{r}
library(here)
source(here("helper","common.R"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial"))
d_sub <- readRDS(here("cached_intermediates","1_d_sub"))
```

Approaches:

- measures [long/short] accuracy, [log] rt, elogit
- across datasets vs. within
- trial level vs. participant level data

Let's not immediately elogit as that was just for models.

Let's also use only `mean(log(rt))` as that's correct if we think RTs are sampled from a log-normal distribution.

# Trial-level, across datasets

Across datasets. Short window looks less wiggly, for whatever reason.

```{r}
d_trial |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

RT

```{r}
d_trial |>
ggplot(aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10() +
  scale_y_log10()
```

# Trial-level, for each dataset

Only look at datasets with some (cross-sectional) variation.

```{r}
datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= 4)

d_trial_agevar <- d_trial |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)
```

```{r}
ggplot(d_trial_agevar, aes(x = age, y = short_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

RT

```{r}
ggplot(d_trial_agevar,
       aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  scale_x_log10() +
  scale_y_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

# Participant-level, across datasets

Accuracy.

```{r}
d_sub |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

RT - but is it the mean of the log or the log of the mean? We choose log of the mean because this fits a log-normal distribution.

```{r}
d_sub |>
ggplot(aes(x = age, y = log_rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10()  + 
  ylab("mean log RT")
```

# Participant-level, for each dataset

Within each dataset.

```{r}
d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)
```

```{r}
ggplot(d_sub_agevar, aes(x = age, y = short_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

elogits

```{r}
ggplot(d_sub_agevar, aes(x = age, y = short_window_elogit)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

RT - here we look at the mean of the log RTs

```{r}
ggplot(d_sub_agevar,
       aes(x = age, y = log_rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  scale_x_log10() +
  scale_y_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

# Appendix: What if we remove zoner trials?

Remove floor and ceiling where there are no shifts?

The answer is that things look even better.

```{r}
d_trial |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
  filter(accuracy > 0, accuracy < 1) |> 
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

```{r}
d_sub_nozoner <- d_trial |>
  filter(short_window_accuracy > 0, short_window_accuracy < 1) |> 
  group_by(dataset_name, administration_id, age) |>
  summarise(across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

```{r}
d_sub_nozoner |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```