### 2024-10-04 2_developmental_analysis.qmd
refactor

---
title: "Trial-level developmental analysis"
format:
  html:
    toc: true
    toc-location: left
---

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(broom.mixed))

source(here("helper","common.R"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial.Rds")) |>
  mutate(log_age = log(age))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
```

Approaches:

- measures [long/short] accuracy, [log] rt, elogit
- across datasets vs. within
- trial level vs. participant level data

Let's not immediately elogit as that was just for models.

Let's also use only `mean(log(rt))` as that's correct if we think RTs are sampled from a log-normal distribution.

TODO:
- number of trials RT cleaning for slope comparison? (25% dataset? 2 RTs)
- check factor invariance across age - Measurement invariance?!?
- report on non-linearity model comparison

# Trial-level

## Across datasets

Across datasets. Short window looks less wiggly, for whatever reason.

```{r}
d_trial |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
  ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

RT

```{r}
d_trial |>
  pivot_longer(c("rt","log_rt"), 
               values_to = "rt", names_to = "measure") |>
  ggplot(aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y")
```
### Stats

Accuracy models.

```{r}
acc_mods <- list(lwe_lin = lm(long_window_elogit ~ age, data = d_trial),
                 lwe_log = lm(long_window_elogit ~ log_age, data = d_trial),
                 swe_lin = lm(short_window_elogit ~ age, data = d_trial),
                 swe_log = lm(short_window_elogit ~ log_age, data = d_trial),
                 lwa_lin = lm(long_window_accuracy ~ age, data = d_trial),
                 lwa_log = lm(long_window_accuracy ~ log_age, data = d_trial),
                 swa_lin = lm(short_window_accuracy ~ age, data = d_trial),
                 swa_log = lm(short_window_accuracy ~ log_age, data = d_trial))

acc_mods_summary <- map_df(acc_mods, ~broom::glance(.x)) |>
  mutate(model = names(acc_mods))

kable(acc_mods_summary, digits = 4)
```

Reaction time models.

```{r}
rt_mods <- list(log_log = lm(log_rt ~ log_age, data = d_trial),
                log_lin = lm(log_rt ~ age, data = d_trial),
                lin_log = lm(rt ~ log_age, data = d_trial),
                lin_lin = lm(rt ~ age, data = d_trial))

rt_mods_summary <- map_df(rt_mods, ~broom::glance(.x)) |>
  mutate(model = names(rt_mods))

kable(rt_mods_summary, digits = 4)
```

## Within dataset

Only look at datasets with some (cross-sectional) variation.

```{r}
datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= 4)

d_trial_agevar <- d_trial |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)
```

Accuracy

```{r}
ggplot(d_trial_agevar, aes(x = age, y = short_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

RT

```{r}
ggplot(d_trial_agevar,
       aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  scale_x_log10() +
  scale_y_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

### Stats

Accuracy models. This time LMERs.

```{r}
d_trial$age_s <- scale(d_trial$age)[,1]
d_trial$log_age_s <- scale(d_trial$log_age)[,1]

acc_mods_lmer <- list(lwe_lin = lmer(long_window_elogit ~ 1 
                                     + (age_s | dataset_name), data = d_trial),
                      lwe_log = lmer(long_window_elogit ~ 1
                                     + (log_age_s | dataset_name), data = d_trial),
                      swe_lin = lmer(short_window_elogit ~ 1
                                     + (age_s | dataset_name), data = d_trial),
                      swe_log = lmer(short_window_elogit ~ 1
                                     + (log_age_s | dataset_name), data = d_trial),
                      lwa_lin = lmer(long_window_accuracy ~ 1
                                     + (age_s | dataset_name), data = d_trial),
                      lwa_log = lmer(long_window_accuracy ~ 1
                                     + (log_age_s | dataset_name), data = d_trial),
                      swa_lin = lmer(short_window_accuracy ~ 1
                                     + (age_s | dataset_name), data = d_trial),
                      swa_log = lmer(short_window_accuracy ~ 1
                                     + (log_age_s | dataset_name), data = d_trial))

acc_mods_lmer_summary <- map_df(acc_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(acc_mods_lmer), 
         r2 = map_dbl(acc_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(acc_mods_lmer_summary, digits = 4)
```

Reaction time models.

```{r}
rt_mods_lmer <- list(log_log = lmer(log_rt ~ 1 + 
                                      (log_age_s | dataset_name), data = d_trial),
                     log_lin = lmer(log_rt ~ 1 + 
                                      (age_s | dataset_name), data = d_trial),
                     lin_log = lmer(rt ~ 1 + 
                                      (log_age_s | dataset_name), data = d_trial),
                     lin_lin = lmer(rt ~ 1 + 
                                      (age_s | dataset_name), data = d_trial))

rt_mods_lmer_summary <- map_df(rt_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_mods_lmer), 
         r2 = map_dbl(rt_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(rt_mods_lmer_summary, digits = 4)
```

# Participant-level

```{r}
d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)
```

## Across datasets

Accuracy.

```{r}
d_sub_long <- d_sub |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") 
d_sub_agevar_long <- d_sub_agevar |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure")

ggplot(d_sub_long, 
       aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long,
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure) + 
  theme(legend.position = "bottom") + 
  ylab("Accuracy") + 
  xlab("Age (months)")
```

Elogits

```{r}
d_sub_long_elogit <- d_sub |>
  pivot_longer(contains("elogit"), 
               values_to = "accuracy", names_to = "measure") 
d_sub_agevar_long_elogit <- d_sub_agevar |>
  pivot_longer(contains("elogit"), 
               values_to = "accuracy", names_to = "measure")

ggplot(d_sub_long_elogit, 
       aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long_elogit,
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure) + 
  theme(legend.position = "bottom") + 
  ylab("Accuracy (elogit)") + 
  xlab("Age (months)")
```

RTs

```{r}
d_sub_long_rt <- d_sub |>
  pivot_longer(c("rt","log_rt"), 
               values_to = "rt", names_to = "measure")

d_sub_agevar_long_rt <- d_sub_agevar |>
  pivot_longer(c("rt","log_rt"), 
               values_to = "rt", names_to = "measure")

ggplot(d_sub_long_rt, 
       aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar_long_rt,
              aes(group = dataset_name, col = dataset_name), geom="line",
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  # geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure, scales = "free_y") + 
  theme(legend.position = "bottom") + 
  ylab("mean log RT") + 
  xlab("Age (months)")

```

### Stats

Accuracy models.

```{r}
acc_mods_sub <- list(lwe_lin = lm(long_window_elogit ~ age, data = d_sub),
                     lwe_log = lm(long_window_elogit ~ log_age, data = d_sub),
                     swe_lin = lm(short_window_elogit ~ age, data = d_sub),
                     swe_log = lm(short_window_elogit ~ log_age, data = d_sub),
                     lwa_lin = lm(long_window_accuracy ~ age, data = d_sub),
                     lwa_log = lm(long_window_accuracy ~ log_age, data = d_sub),
                     swa_lin = lm(short_window_accuracy ~ age, data = d_sub),
                     swa_log = lm(short_window_accuracy ~ log_age, data = d_sub))

acc_mods_sub_summary <- map_df(acc_mods_sub, ~broom::glance(.x)) |>
  mutate(model = names(acc_mods_sub))

kable(acc_mods_sub_summary, digits = 4)
```

Reaction time models.

```{r}
mods_sub <- list(log_log = lm(log_rt ~ log_age, data = d_sub),
                 log_lin = lm(log_rt ~ age, data = d_sub),
                 lin_log = lm(rt ~ log_age, data = d_sub),
                 lin_lin = lm(rt ~ age, data = d_sub))

models_sub <- map_df(mods_sub, ~broom::glance(.x)) |>
  mutate(model = names(mods_sub))

kable(models_sub, digits = 4)
```

## Within dataset

Within each dataset.

```{r}
ggplot(d_sub_agevar, aes(x = age, y = long_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

elogits

```{r}
ggplot(d_sub_agevar, aes(x = age, y = long_window_elogit)) +
  geom_point(alpha = .05) +
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  geom_hline(yintercept = .5, lty = 2) +
  # scale_x_log10() +
  ylim(-2,3) + 
  facet_wrap(~dataset_name, scales = "free_x")
```

RT - here we look at the mean of the log RTs

```{r}
ggplot(d_sub_agevar,
       aes(x = age, y = log_rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth(method = "lm", col = "green", alpha = .5) +
  geom_smooth(method = "loess", alpha = .5) +
  scale_x_log10() +
  scale_y_log10() + 
  facet_wrap(~dataset_name, scales = "free_x")
```

### Stats

Accuracy models.

```{r}
d_sub$age_s <- scale(d_sub$age)[,1]
d_sub$log_age_s <- scale(d_sub$log_age)[,1]

acc_mods_sub_lmer <- list(lwe_lin = lmer(long_window_elogit ~ 1 
                                         + (age_s | dataset_name), data = d_sub),
                          lwe_log = lmer(long_window_elogit ~ 1
                                         + (log_age_s | dataset_name), data = d_sub),
                          swe_lin = lmer(short_window_elogit ~ 1
                                         + (age_s | dataset_name), data = d_sub),
                          swe_log = lmer(short_window_elogit ~ 1
                                         + (log_age_s | dataset_name), data = d_sub),
                          lwa_lin = lmer(long_window_accuracy ~ 1
                                         + (age_s | dataset_name), data = d_sub),
                          lwa_log = lmer(long_window_accuracy ~ 1
                                         + (log_age_s | dataset_name), data = d_sub),
                          swa_lin = lmer(short_window_accuracy ~ 1
                                         + (age_s | dataset_name), data = d_sub),
                          swa_log = lmer(short_window_accuracy ~ 1
                                         + (log_age_s | dataset_name), data = d_sub))

acc_mods_sub_lmer_summary <- map_df(acc_mods_sub_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(acc_mods_sub_lmer), 
         r2 = map_dbl(acc_mods_sub_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(acc_mods_sub_lmer_summary, digits = 4)
```
With fixed effects.

```{r}
acc_mods_sub_lmer_fe <- list(lwe_lin = lmer(long_window_elogit ~ age_s  
                                         + (age_s | dataset_name), data = d_sub),
                          lwe_log = lmer(long_window_elogit ~ log_age_s 
                                         + (log_age_s | dataset_name), data = d_sub),
                          swe_lin = lmer(short_window_elogit ~ age_s 
                                         + (age_s | dataset_name), data = d_sub),
                          swe_log = lmer(short_window_elogit ~ log_age_s 
                                         + (log_age_s | dataset_name), data = d_sub),
                          lwa_lin = lmer(long_window_accuracy ~ age_s
                                         + (age_s | dataset_name), data = d_sub),
                          lwa_log = lmer(long_window_accuracy ~ log_age_s
                                         + (log_age_s | dataset_name), data = d_sub),
                          swa_lin = lmer(short_window_accuracy ~ age_s
                                         + (age_s | dataset_name), data = d_sub),
                          swa_log = lmer(short_window_accuracy ~ log_age_s
                                         + (log_age_s | dataset_name), data = d_sub))

acc_mods_sub_lmer_summary_fe <- map_df(acc_mods_sub_lmer_fe, ~broom.mixed::glance(.x)) |>
  mutate(model = names(acc_mods_sub_lmer_fe), 
         r2 = map_dbl(acc_mods_sub_lmer_fe, 
                      ~performance::r2_nakagawa(.x)$R2_conditional))

kable(acc_mods_sub_lmer_summary_fe, digits = 4)
```

### Reaction time models.

```{r}
rt_mods_lmer <- list(log_log = lmer(log_rt ~ 1 + 
                                      (log_age_s | dataset_name), data = d_sub),
                     log_lin = lmer(log_rt ~ 1 + 
                                      (age_s | dataset_name), data = d_sub),
                     lin_log = lmer(rt ~ 1 + 
                                      (log_age_s | dataset_name), data = d_sub),
                     lin_lin = lmer(rt ~ 1 + 
                                      (age_s | dataset_name), data = d_sub))

rt_mods_lmer_summary <- map_df(rt_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_mods_lmer), 
         r2 = map_dbl(rt_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(rt_mods_lmer_summary, digits = 4)
```

```{r}
rt_mods_lmer_fe <- list(log_log = lmer(log_rt ~ log_age_s + 
                                      (log_age_s | dataset_name), data = d_sub),
                     log_lin = lmer(log_rt ~ age_s + 
                                      (age_s | dataset_name), data = d_sub),
                     lin_log = lmer(rt ~ log_age_s + 
                                      (log_age_s | dataset_name), data = d_sub),
                     lin_lin = lmer(rt ~ age_s + 
                                      (age_s | dataset_name), data = d_sub))

rt_mods_lmer_summary_fe <- map_df(rt_mods_lmer_fe, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_mods_lmer_fe), 
         r2 = map_dbl(rt_mods_lmer_fe, ~performance::r2_nakagawa(.x)$R2_conditional))

kable(rt_mods_lmer_summary_fe, digits = 4)
```

# Appendix: What if we remove zoner trials?

Remove floor and ceiling where there are no shifts?

The answer is that things look even better.

```{r}
d_trial |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
  filter(accuracy > 0, accuracy < 1) |> 
  ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

```{r}
d_sub_nozoner <- d_trial |>
  filter(short_window_accuracy > 0, short_window_accuracy < 1) |> 
  group_by(dataset_name, administration_id, age) |>
  summarise(across(c(contains("_window"),"rt","log_rt"), ~mean(.x, na.rm=TRUE)))
```

```{r}
d_sub_nozoner |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
  ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```