### 2024-10-28 6_paper_analyses.qmd

---
title: "Analyses in the paper"
format: html
---

```{r}

suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tidySEM))

source(here("helper","common.R"))
d_trial <- readRDS(here("cached_intermediates","1_d_trial.Rds")) |>
  mutate(log_age = log(age))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))

d_trial$age_s <- scale(d_trial$age)[,1]
d_trial$log_age_s <- scale(d_trial$log_age)[,1]

```

# Datasets table

```{r}
d_sub |>
  group_by(dataset_name) |>
  summarise(`N subjects` = length(unique(subject_id)),
            `N admins` = n(), 
            `Mean Age` = mean(age), 
            `Min Age` = min(age), 
            `Max Age` = max(age), 
            `Avg Trials` = mean(n_trials),
            `Avg RT Trials` = mean(n_trials_rt),
            `CDIs` = ifelse(!is.na(any(prod)), "x", ""), 
            `longitudinal` = ifelse(any(duplicated(subject_id)), "x", "")) |>
  arrange(desc(`N subjects`)) |>
  xtable(caption = "Characteristics of included datasets from Peekbank. `Admins` denotes separate experimental sessions.")
```

# Speed and accuracy

Accuracy

```{r}
acc_mod = lmer(long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = d_trial)

summary(acc_mod)
```

Reaction time.

```{r}
rt_mod = lmer(rt ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = d_trial)

summary(rt_mod)
```

```{r}

datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= AGEVAR_CUTOFF)

d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)

a <- ggplot(d_sub,
       aes(x = age, y = long_window_accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  ylab("Accuracy") + 
  xlab("Age (months)")

b <- ggplot(d_sub,
       aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = .1, method = "lm", size = .5, lty = 2) + 
  # geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  scale_color_discrete(guide = FALSE) + 
  theme(legend.position = "bottom") + 
  ylab("Reaction Time (log)") + 
  xlab("Age (months)")

f <- cowplot::plot_grid(a,b)
f
ggsave(f, filename = "figures/acc_rt.png", width = 8, height = 3.5)

```

# Variability

```{r}
d_var <- d_sub |>
  rename(RT = rt_var,
         Accuracy = long_window_acc_var) |>
  pivot_longer(c("RT","Accuracy"), 
               values_to = "sd", names_to = "measure")

v <- ggplot(d_var, aes(x = age, y = sd)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_smooth(data = d_var |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name),
              aes(group = dataset_name, col = dataset_name), 
              alpha = .1, method = "lm", size = .5, lty = 2, se=FALSE) + 
  scale_color_discrete(guide = FALSE) + 
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y") + 
  ylab("Standard Deviation") + 
  xlab("Age (months)")

v 
ggsave(v, filename = "figures/var.png", width = 8, height = 3.5)

```

Accuracy

```{r}
d_var$age_s <- scale(d_var$age)[,1]
d_var$log_age_s <- scale(d_var$log_age)[,1]

acc_var_mod <- lmer(sd ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = filter(d_var, measure == "Accuracy"))

summary(acc_mod)
```

Reaction time.

```{r}
rt_var_mod <- lmer(sd ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = filter(d_var, measure == "RT"))
summary(rt_var_mod)
```

# Factor analysis

```{r}
d_cfa <- d_sub |>
  rename(`accuracy` = long_window_accuracy) |>
  mutate(accuracy = scale(accuracy), 
         log_age = scale(log_age), 
         rt = scale(rt)) |>
  select(dataset_name, accuracy, log_age, rt)

d_cfa <- d_cfa[complete.cases(d_cfa), ]

fa_model <-  "F1  =~ log_age + rt + accuracy"
fit <- cfa(fa_model, d_cfa, std.lv=TRUE)

summary(fit)
graph_sem(model = fit)+ theme(panel.background = element_rect(fill = "white"))
```