### 2025-03-13 paper/paper.Rmd
nilam

---
title: Continuous developmental changes in word recognition support language learning across early childhood

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author:
  - name: Michael C. Frank
    affiliation: a, 1
  - name: Virginia A. Marchman
    affiliation: a
  - name: Claire Augusta Bergey
    affiliation: b
  - name: Veronica Boyce
    affiliation: a
  - name: Mika Braginsky
    affiliation: a
  - name: Jess Mankewitz
    affiliation: c
  - name: Stephan Meylan
    affiliation: d
  - name: Ben Prystawski
    affiliation: a
  - name: Nilam Ram
    affiliation: a, e
  - name: Robert Z. Sparks
    affiliation: a
  - name: Adrian Steffan
    affiliation: f
  - name: Alvin Wei Ming Tan
    affiliation: a
  - name: Martin Zettersten
    affiliation: f
address:
  - code: a
    address: Department of Psychology, Stanford University, Stanford, CA, USA
  - code: b
    address: Department of Linguistics, Stanford University, Stanford, CA, USA
  - code: c
    address: Department of Psychology, University of Wisconsin, Madison, WI, USA
  - code: d
    address: Department of Linguistics, University of California, Berkeley, CA, USA
  - code: e
    address: Department of Communication, Stanford University, Stanford, CA, USA
  - code: f
    address: Department of Communication, Stanford University, Stanford, CA, USA
  - code: g
    address: Department of Psychology, Ludwig Maximilian University, Munich, Germany

corresponding_author:
  code: 1
  text: "To whom correspondence should be addressed. E-mail: mcfrank@stanford.edu"

# For footer text
lead_author_surname: Anonymous

author_contributions: |
  Please provide details of author contributions here.

abstract: |
  Being a fluent language user involves recognizing words as they unfold in time. How does this skill develop over the course of early childhood?  And how does facility in word recognition relate to the growth of vocabulary knowledge?  We address these questions using data from Peekbank, an open database of experiments measuring children's eye-movements during early word recognition.  Combining 24 datasets from almost 2,000 children ages 1--6 years, we show that word recognition becomes faster, more accurate, and less variable across development. Factor analysis reveals a cross-sectional coupling of word recognition speed and accuracy with parent-reported vocabulary. Across a range of models, speed, accuracy, and vocabulary also show coupled growth such that children with faster initial word recognition tend to show faster vocabulary growth. Together, these findings support the view that word recognition is a skill that develops gradually across early childhood and that this skill plays a role in supporting early language learning.

significance: |
  The efficiency with which children recognize words provides a measurement of their real-time language processing, which has been argued to be a key part of the language learning process. Here, we use a large dataset of eye-tracking data from many different experiments to map out the development of word recognition, finding that this process becomes faster, more accurate, and less variable in the period between one and six years of age, and that children who are better at word recognition tend to have larger and faster growing vocabularies. These data suggest that understanding language is a skill that improves gradually with practice.

acknowledgements: |
  This work was supported by a grant from the Jacobs Foundation.

keywords:
  - language development
  - word recognition
  - developmental change
  - skill acquisition
  - eye-tracking

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: peekbank.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

output: rticles::pnas_article
---

\dropcap{C}hildren acquiring a language are learning a body of knowledge -- a set of words and the ways they are combined -- but they are also learning to deploy this knowledge in the myriad complex, noisy, and fast-moving environments in which language is used. As children enter their second year, language explodes onto the scene; both vocabulary and grammatical abilities grow rapidly and in tandem [@bates1994developmental;@frank2021]. This growth in knowledge is also accompanied by changes in language processing efficiency: children become quicker and more accurate in recognizing words and matching them with their referents [@fernald1998;@peter2019;@bergelson2020comprehension].

Yet unlike language production, which is manifest via overt behavior, evidence for word recognition is often more subtle. Very young children with incomplete knowledge may not be able to point to the correct referent of a word, but they may still have some representation of word meaning [@bergelson2012]. Eye tracking has thus emerged as a key method that allows the measurement of language comprehension with high temporal resolution: both adults and children reliably fixate the referent of a word soon after it is used [@tanenhaus1995integration;@altmann1999incremental;@fernald1998;@fernald2008;@macdonald2018]. The relative timecourse of fixation then can provide an index of the comprehender's ability or the difference between two stimulus conditions.

The version of this method that is used with children goes by many names, including the intermodal preferential looking paradigm and the "looking while listening" paradigm (LWL, the name we adopt here) [@hirsh1996intermodal;@reznick1990visual;@fernald2008]. In LWL experiments, children are typically shown a series of trials in which two images are displayed side by side and they are asked to find one of them. For example, a ball and a book might be shown, and the child might hear "Look at the ball! Can you find it?". Accuracy is then computed as the proportion of time they fixate the correct image within a fixed window after the onset of the noun ("ball" in this case). Reaction time is computed only on trials in which the child is fixating on the distractor image (the book) before word onset; in these cases, the average time it takes until the child shifts fixation from the distractor to the target is used as an index of processing speed. Early work using this method showed that both children's speed and accuracy increase rapidly across the second year [@fernald1998;@reznick1990visual]. Related methods have provided a window into how children process phonological [@mani2011], syntactic [@trueswell1999kindergarten], and semantic [@mani2010;@bergelson2017] information as well as how their lexical representations develop [@Swingley2002].

Word recognition ability, as measured by LWL, is hypothesized to play a key role in language learning. Each word that a child experiences is an opportunity to learn; measurements of children's language input at home are consistently associated with their vocabulary size [@hart1995meaningful;@anderson2021linking]. Recognizing incoming words and linking them with their referents is a prerequisite for learning. Consider a child hearing the utterance "Can you put the ball in the box?" The faster and more accurately the child can recognize that the ball is a referent, the better they can use this evidence to help infer the speaker's intended meaning, perhaps making inferences about the meaning of "put" or "box" [@frank2009using]. Consistent with this idea, one important study found that children's word recognition speed mediated the longitudinal relationship between home language input and vocabulary growth [@Weisleder2013].

\begin{table*}[ht]
\centering
\begin{tabular}{rlrrrrrrrll}
  \hline
 & dataset\_name & N subjects & N admins & Mean Age & Min Age & Max Age & Avg Trials & Avg RT Trials & CDIs & longitudinal \\ 
  \hline
1 & reflook\_v4 & 310 & 310 & 36.82 & 12.24 & 60.00 & 6.21 & 2.87 &  &  \\ 
  2 & yurovsky\_2017 & 282 & 282 & 25.64 & 12.59 & 58.65 & 5.95 & 2.79 &  &  \\ 
  3 & weaver\_zettersten\_2024 & 141 & 248 & 15.73 & 13.50 & 23.60 & 18.10 & 6.72 &  & x \\ 
  4 & fernald\_marchman\_2012 & 122 & 678 & 23.92 & 17.00 & 32.00 & 16.89 & 7.21 & x & x \\ 
  5 & sander-montant\_2022 & 122 & 122 & 21.94 & 12.02 & 31.11 & 10.07 & 4.20 & x &  \\ 
  6 & frank\_tablet\_2016 & 104 & 104 & 33.89 & 12.13 & 59.84 & 5.87 & 2.74 &  &  \\ 
  7 & baumgartner\_2014 & 100 & 100 & 12.01 & 12.00 & 13.00 & 4.00 & 2.31 & x &  \\ 
  8 & fmw\_2013 &  80 & 179 & 20.04 & 17.00 & 26.00 & 21.26 & 8.44 & x & x \\ 
  9 & adams\_marchman\_2018 &  69 & 711 & 23.58 & 13.00 & 38.00 & 18.44 & 7.92 & x & x \\ 
  10 & potter\_canine &  67 &  67 & 23.76 & 21.00 & 27.00 & 10.38 & 3.80 & x &  \\ 
  11 & fernald\_totlot &  63 & 229 & 19.68 & 15.00 & 25.00 & 15.25 & 6.06 & x & x \\ 
  12 & pomper\_prime &  63 &  63 & 39.73 & 37.90 & 42.00 & 12.98 & 5.34 &  &  \\ 
  13 & pomper\_saffran\_2016 &  60 &  60 & 44.27 & 41.00 & 47.00 & 7.55 & 3.30 &  &  \\ 
  14 & swingley\_aslin\_2002 &  50 &  50 & 15.09 & 14.13 & 16.00 & 11.74 & 3.79 & x &  \\ 
  15 & pomper\_salientme &  44 &  44 & 40.11 & 38.00 & 43.00 & 5.30 & 2.34 &  &  \\ 
  16 & perry\_cowpig &  42 &  42 & 20.45 & 19.00 & 22.00 & 14.88 & 5.43 & x &  \\ 
  17 & ronfard\_2021 &  40 &  40 & 19.95 & 18.00 & 24.00 & 18.54 & 7.62 & x &  \\ 
  18 & bacon\_gendercues &  38 &  38 & 22.87 & 22.00 & 24.00 & 18.16 & 8.00 & x &  \\ 
  19 & garrison\_bergelson\_2020 &  35 &  35 & 14.46 & 12.00 & 18.00 & 27.47 & 9.41 & x &  \\ 
  20 & pomper\_yumme &  32 &  32 & 26.38 & 25.00 & 28.00 & 7.62 & 2.31 &  &  \\ 
  21 & mahr\_coartic &  29 &  29 & 20.83 & 18.10 & 23.80 & 24.38 & 8.86 & x &  \\ 
  22 & ferguson\_eyetrackingr &  28 &  28 & 19.71 & 18.02 & 21.86 & 5.54 & 2.33 & x &  \\ 
  23 & potter\_remix &  23 &  44 & 22.59 & 18.00 & 29.00 & 6.07 & 2.58 & x & x \\ 
  24 & newman\_genderdistractor &  19 &  19 & 30.12 & 29.23 & 30.84 & 13.11 & 4.94 &  &  \\ 
  \hline
    & Total & 1963 & 3554 & 24.73 & 12.00 & 60.00 & 12.74 & 5.05 & 15 & 6 \\ 
   \hline
\end{tabular}
\caption{Characteristics of included datasets from Peekbank. `Admins` denotes separate experimental sessions. `CDIs` refers to whether the dataset contains parent report vocabulary data from the MacArthur-Bates Communicative Development Inventory.} 
\end{table*}

Word recognition speed has been used as an index of individual differences in early childhood [@fernald2006;@marchman2008;@peter2019;@fernald2012;@newbury2016interrelationships] and beyond [@colby2023efficiency;@jeppsen2022development;@mcmurray2022slow]. Over and above measures of vocabulary size, word recognition speed at 18 months predicts children's standardized test scores years later [@marchman2008], though these predictions may be limited to particular ages or processing assessments [@peter2019]. Further, faster processing at 18 months is predictive of whether "late talkers" catch up to their peers or require further intervention [@fernald2012].  Critically, all of these assessments use words that children are reported to understand and produce -- they are not indices of vocabulary size but rather of how quickly and accurately they can recognize a spoken word and use it to guide their visual attention to a referent. 
Yet individual experiments measuring processing with young children typically recruit relatively small samples in a restricted range of ages. These samples provide neither the breadth of ages nor the number of participants needed to estimate the broader dynamics of how word recognition changes developmentally and how it connects with other aspects of early language development (see [@colby2023efficiency; @mcmurray2022slow] for examples of these analyses in school-aged children). To overcome this limitation, we turned to Peekbank, an open database of LWL data from young children, stored in a harmonized format [@zettersten2023peekbank]. This resource allows us to gain an unprecedented view of the development of word recognition across a large sample of children.

Here we investigate two specific hypotheses. First, one influential theory characterizes language learning as a process of learning the skill of real-time processing [@christiansen2016;@chater2018]. Based on this theory, we should expect to see the signatures of expertise and skill learning in word recognition. Accuracy should change linearly with the logarithm of age, reflecting gradual asymptotic convergence to mature levels of accuracy. In addition, the logarithm of processing speed should decrease with the logarithm of age, potentially reflecting the "power law of practice" [@snoddy1926;@kail1991;@anderson1982; cf @heathcote2000]. Finally, trial-to-trial variability in both speed and accuracy should decrease with increasing expertise [@todorov2002optimal].

Second, previous findings have provided limited and sometimes conflicting evidence on the concurrent and predictive relations between word recognition and language learning. Initial reports showed strong predictive relationships between both speed and accuracy and later vocabulary growth [@fernald2006], with replications in infants born preterm [@marchman2016] and late talkers [@fernald2012]. Subsequent studies have primarily focused on speed of processing and found more mixed results, with reaction time measures found to be only inconsistently predictive of later vocabulary outcomes [@peter2019;@newbury2016interrelationships;@lany2018]. A larger dataset should allow us to make a more definitive test of the presence of these relationships. Further, by examining the relationship between speed, accuracy, and vocabulary, it should be possible to assess the extent to which processing speed specifically plays a role in vocabulary growth.

```{r loading, echo=FALSE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(tidySEM))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(papaja))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(nlme))

source(here("helper","common.R"))

source(here("helper/modified_sem_plot.R"))

d_trial <- readRDS(here("cached_intermediates","1_d_trial.Rds")) |>
  mutate(log_age = log(age))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds")) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(admin_num = 1:n(), 
         time_since_t0 = age - age[1],
         delta_t = c(0, diff(age)))

d_trial$age_s <- scale(d_trial$age)[,1]
d_trial$log_age_s <- scale(d_trial$log_age)[,1]   
d_sub$age_s <- scale(d_sub$age)[,1]
d_sub$log_age_s <- scale(d_sub$log_age)[,1]

AGEVAR_CUTOFF <- 6

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
```

# Results {.unnumbered}

We retrieved data from Peekbank, focusing on data from children ages 1--6 years. Although experiments in Peekbank include a variety of different experimental manipulations, we analyzed data only from simple word recognition trials in which children were shown two pictures of concrete objects and heard a label for an object (typically embedded in a simple carrier phrase such as "Look at the ..."); these trials often constituted control conditions for experiments. We focus here on English purely for practical reasons -- the Peekbank dataset contains limited data from other languages.

These criteria yielded `r length(unique(d_sub$dataset_name))` datasets, including `r length(unique(d_sub$subject_id))` children and `r length(unique(d_sub$administration_id))` administrations of the LWL procedure (some datasets were longitudinal or involved multiple closely-spaced testing sessions). Table \ref{tab:datasets} shows the characteristics of individual datasets (see also Figures S1 and S2). The size of the combined dataset, the unified data processing pipeline, and the fact that individual studies used very similar implementations of the LWL experimental paradigm, all allowed us to make a more detailed study of the development of word recognition than has previously been possible. While our analysis is exploratory in nature, it is guided by the two hypotheses outlined above: the presence of 1) signatures of skill learning in word recognition, and 2) linkages between word recognition and vocabulary.

```{r dataset_table}
dataset_characteristics <- d_sub |>
  group_by(dataset_name) |>
  summarise(`N subjects` = length(unique(subject_id)),
            `N admins` = n(), 
            `Mean Age` = mean(age), 
            `Min Age` = min(age), 
            `Max Age` = max(age), 
            `Avg Trials` = mean(n_trials[!is.na(long_window_accuracy)]),
            `Avg RT Trials` = mean(n_trials_rt[!is.na(rt)]),
            `CDIs` = ifelse(!is.na(any(prod)), "x", ""), 
            `longitudinal` = ifelse(any(duplicated(subject_id)), "x", "")) |>
  arrange(desc(`N subjects`))

dataset_characteristics <- dataset_characteristics |>
  bind_rows(tibble(dataset_name = "Total", 
                       `N subjects` = sum(dataset_characteristics$`N subjects`),
                       `N admins` = sum(dataset_characteristics$`N admins`), 
                       `Mean Age` = mean(dataset_characteristics$`Mean Age`),
                       `Min Age` = min(dataset_characteristics$`Min Age`), 
                       `Max Age` = max(dataset_characteristics$`Max Age`), 
                       `Avg Trials` = mean(dataset_characteristics$`Avg Trials`),
                       `Avg RT Trials` = mean(dataset_characteristics$`Avg RT Trials`),
                       `CDIs` = as.character(sum(dataset_characteristics$CDIs != "")),
                       `longitudinal` = as.character(sum(dataset_characteristics$longitudinal != ""))))
```

```{r eval=FALSE}
xtable(dataset_characteristics, 
       caption = "Characteristics of included datasets from Peekbank. `Admins` denotes separate experimental sessions.")
```

## Speed and accuracy of word recognition increase {.unnumbered}

```{r acc_mods}
d_trial$age_s <- scale(d_trial$age)[,1]
d_trial$log_age_s <- scale(d_trial$log_age)[,1]

acc_mods_lmer <- list(lwa_lin = lmer(long_window_accuracy ~ age_s  +
                                     + (age_s | dataset_name) + (1 | subject_id), data = d_trial),
                      lwa_log = lmer(long_window_accuracy ~ log_age_s +
                                     + (log_age_s | dataset_name) + (1 | subject_id), data = d_trial),
                      swa_lin = lmer(short_window_accuracy ~ age_s  +
                                     + (age_s | dataset_name) + (1 | subject_id), data = d_trial),
                      swa_log = lmer(short_window_accuracy ~ log_age_s +
                                     + (log_age_s | dataset_name) + (1 | subject_id), data = d_trial))

acc_mods_lmer_summary <- map_df(acc_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(acc_mods_lmer), 
         r2 = map_dbl(acc_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))
```

```{r rt_mods}
rt_mods_lmer <- list(log_lin = lmer(log_rt ~ age_s + 
                                      (age_s | dataset_name) + (1 | subject_id), data = d_trial),
                     log_log = lmer(log_rt ~ log_age_s + 
                                      (log_age_s | dataset_name) + (1 | subject_id), data = d_trial),
                     lin_lin = lmer(rt ~ age_s + 
                                      (age_s | dataset_name) + (1 | subject_id), data = d_trial),
                     lin_log = lmer(rt ~ log_age_s +  
                                      (log_age_s | dataset_name) + (1 | subject_id), data = d_trial))

rt_mods_lmer_summary <- map_df(rt_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_mods_lmer), 
         r2 = map_dbl(rt_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))
```

We began by characterizing developmental changes in speed and accuracy. We computed both RTs (reaction times) and accuracies following standard practices in the literature [@fernald2008]. Because there is no consensus about the length of time windows for the computation of accuracy, we considered both a shorter window (from 200 -- 2200 ms after noun onset) and a longer window (from 200 -- 4000 ms). For each window, we averaged all fixation within the window to compute a continuous proportion between 0 (no fixation on the target during the window) and 1 (total fixation on the target during the window) on every trial.

Our first question was about the functional form of the relationships between age, speed, and accuracy (see SI for raw correlations between variables). To investigate this question, we fit linear mixed effects models predicting accuracy and RT on each trial across the full dataset with random slopes of child age by study and random intercepts by participant. We compared models that included both long and short accuracy windows, as well as logarithmic and linear effects of age, and logarithmic and linear transformations of RT. The best fitting model of accuracy predicted long window accuracy as a function of the logarithm of age,^[This longer window yielded overall higher cross-trial reliability as well, and so we report long window results in the main text. All findings reported here hold for both window sizes, however (see SI).] and the best fitting model of speed predicted log RT as a function of log age as well (see SI).

Figure \ref{fig:devchange} shows these age gradients. Log RT decreased significantly with age (`r papaja::apa_print(rt_mods_lmer[[2]])$full_result$log_age_s`) and accuracy increased significantly with age  (`r papaja::apa_print(acc_mods_lmer[[2]])$full_result$log_age_s`). In sum, we see continuing improvements in word recognition across the full age range in our dataset that appear roughly linear in the logarithm of age.  These logarithmic relationships follow theoretical expectations that both speed and accuracy should gradually asymptote to mature levels of performance as in skill learning more generally [@anderson1982,@snoddy1926]. 
<!-- With our current dataset we cannot specifically determine whether within-individual patterns of change conform to power law or exponential forms [@kail1991,@heathcote2000], but regardless we see clear patterns of developmental change related to skill acquisition. -->

```{r devchange, fig.cap="Participant-level accuracy and reaction time (log), plotted by age (log). The solid blue line shows a linear fit and associated confidence interval. Individual dotted lines show linear fits for those datasets spanning six or more months of age. The dashed line for accuracy shows chance-level looking (.5)", fig.env="figure*", fig.height = 2.5}

datasets_with_age_variation <- d_trial |>
  group_by(dataset_name) |>
  summarise(age_range = max(age) - min(age)) |>
  filter(age_range >= AGEVAR_CUTOFF)

d_sub_agevar <- d_sub |>
  filter(dataset_name %in% datasets_with_age_variation$dataset_name)

#library(RColorBrewer)

  mycolors = c("#1B9E77", "#D95F02", "#E7298A", "#66A61E", "#1B9E77", "#D95F02", "#E7298A", "#66A61E", "#1B9E77", "#D95F02", "#E7298A", "#66A61E","#1B9E77", "#D95F02")
  # 4 colors from brewer scale "Dark2" that I thought had decent contrast against the grey/black background dots

a <- ggplot(d_sub,
       aes(x = age, y = long_window_accuracy)) + 
  geom_point(alpha = .02) + 
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = 1, method = "lm", size = .5, lty = 3) + 
    geom_smooth(method = "lm") +
  geom_hline(yintercept = .5, lty = 3) + 
  scale_x_log10(breaks = c(12,24,36,48,60)) +
  scale_color_manual(values=mycolors, guide = FALSE) + 
  ylab("Accuracy") + 
  xlab("Age (months)")

b <- ggplot(d_sub,
       aes(x = age, y = rt)) + 
  geom_point(alpha = .02) + 
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = .1, method = "lm", size = .5, lty = 3) + 
    geom_smooth(method = "lm") +
  scale_x_log10(breaks = c(12,24,36,48,60)) +
  scale_y_log10() +
  scale_color_manual(values=mycolors, guide = FALSE) + 
  theme(legend.position = "bottom") + 
  ylab("Reaction Time (log)") + 
  xlab("Age (months)")

cowplot::plot_grid(a,b)
```

## Variability of word recognition decreases {.unnumbered}

```{r var_mods}
acc_var_mod <- lmer(long_window_acc_var ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = d_sub)

rt_var_mod <- lmer(log_rt_var ~ log_age_s + (log_age_s | dataset_name) + 
                 (1 | subject_id), 
               data = d_sub)
```

One further hallmark of increasing skill is a decrease in task-relevant variability [@todorov2002optimal]. Both within and across datasets, within-individual variation in accuracy and RT decreased across the developmental range we examined (Figure \ref{fig:variance}). We fit mixed effects models to the standard deviation of both accuracy and RT for each testing session for each participant, including random slopes of log age by dataset and random intercepts for each participant. For both speed and accuracy, within-individual variability decreased with age (RT: `r papaja::apa_print(rt_var_mod)$full_result$log_age_s`; accuracy: `r papaja::apa_print(acc_var_mod)$full_result$log_age_s`). Thus, as well as being faster and more accurate, older children were more consistent in their online word recognition than younger children.

```{r variance, fig.cap="Participant-level variability in accuracy and reaction time (log RT), plotted by age (log). Plotting conventions are as in Figure 1.", fig.env="figure*", fig.height = 2.5}
  mycolors = c("#1B9E77", "#D95F02", "#E7298A", "#66A61E", "#1B9E77", "#D95F02", "#E7298A", "#66A61E", "#1B9E77", "#D95F02", "#E7298A", "#66A61E","#1B9E77", "#D95F02")

c <- ggplot(d_sub,
       aes(x = age, y = long_window_acc_var)) + 
  geom_point(alpha = .02) + 
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = .1, method = "lm", size = .5, lty = 3) + 
    geom_smooth(method = "lm") +
  scale_x_log10(breaks = c(12,24,36,48,60)) +
  scale_color_manual(values=mycolors, guide = FALSE) + 
  ylab("SD of Accuracy") + 
  xlab("Age (months)")

d <- ggplot(d_sub,
       aes(x = age, y = log_rt_var)) + 
  geom_point(alpha = .02) + 
  geom_smooth(data = d_sub_agevar,
              aes(group = dataset_name, col = dataset_name), se=FALSE,
              alpha = .1, method = "lm", size = .5, lty = 3) + 
    geom_smooth(method = "lm") +
  scale_x_log10(breaks = c(12,24,36,48,60)) +
  scale_color_manual(values=mycolors, guide = FALSE) + 
  theme(legend.position = "bottom") + 
  ylab("SD of log RT") + 
  xlab("Age (months)")

cowplot::plot_grid(c,d)
```

## Speed and accuracy relate to vocabulary size {.unnumbered}

We were next interested in whether the various aspects of word recognition -- including accuracy, RT, and the variability of each of these -- were related to other aspects of early language ability. Of the studies in our database, `r sum(str_length(dataset_characteristics$CDIs)>0)` gathered parent reports about children's early vocabulary using the MacArthur-Bates Communicative Development Inventory (CDI), a popular survey instrument that provides a reliable and valid estimate of children's early vocabulary [@marchman2023;@frank2021]. Different forms of the CDI can be used to measure either receptive and expressive vocabulary (for children up to 18 months) or expressive vocabulary only (for children 16 -- 30 months).

```{r efa}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(dataset_name, log_rt, log_rt_var, long_window_accuracy, long_window_acc_var, prod, comp, age, log_age)

d_sub_mat_s <- d_sub_mat |>
  ungroup() |>
  mutate(across(all_of(c("log_rt", "log_rt_var", "long_window_accuracy", 
                         "long_window_acc_var", "prod", "comp")), 
                       ~ age_scale(.x, age)))

parallel_ana <- capture.output(fa.parallel(select(d_sub_mat, -dataset_name, -age, -log_age), fa = "fa", 
            use = "pairwise.complete.obs", plot=FALSE))
```

```{r cfa}
fa3_model <- "vocab =~ prod + comp
              accuracy =~ long_window_accuracy + long_window_acc_var
              speed =~ log_rt + log_rt_var"

fit3 <- cfa(fa3_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

fit_stats <- suppressMessages(summary(fit3, fit.measures=TRUE, standardize=TRUE))
```

We fit a series of factor analytic models to explore the dimensionality of the parent report and child LWL data. Initial exploratory factor analysis using parallel analysis to select the number of factors suggested that three factors explained substantial variance in the data (see SI: Factor Analysis). To better accommodate missing data under the assumption of data missing at random (e.g., due to dataset), we used confirmatory factor analysis with full information maximum likelihood to find the best set of loadings. The best fitting model of those we examined was a three-factor model with factors for speed (RT and RT variability), accuracy (accuracy and accuracy variability), and vocabulary (comprehension and production from the CDI). Fit statistics for this model were generally good (Confirmatory fit index: `r round(fit_stats$fit["cfi"], 3)`, RMSE: `r round(fit_stats$fit["rmsea"], 3)`); see SI: Alternative Factor Structures).

```{r age_regression, fig.env="figure*", fig.cap="Structural equation model showing the three-factor factor analysis with a regression of each latent variable on the logarithm of age. Observed variables are notated as squares and latent variables are notated as circles Factor loadings and regression coefficients are shown with straight, solid lines; covariances are shown with dashed lines; residual variances are shown as solid circular connections. Stars show conventional levels of statistical significance, e.g. * indicates p < .05, ** indicates p < .01, and *** indicates p < .001. Covariances reflect age-residualized correlations between variables. Abbreviations: acc = accuracy, prod = production vocabulary, comp = comprehension vocabulary.", fig.height = 3}

d_sub_mat_s_renamed <- d_sub_mat_s |>
  rename(acc = long_window_accuracy, 
         acc_sd = long_window_acc_var, 
         log_rt_sd = log_rt_var)

fa3_age_model <- "
# measurement model
vocab =~ prod + comp
accuracy =~ acc + acc_sd
speed =~ log_rt + log_rt_sd

# regressions
vocab ~ log_age
accuracy ~ log_age
speed ~ log_age
"

fit3_age <- cfa(fa3_age_model, d_sub_mat_s_renamed, std.lv=TRUE, missing='fiml')

fit3_age_summary <- suppressMessages(summary(fit3_age, fit.measures=TRUE, standardize=TRUE))
layout_age = matrix(nrow=5, ncol = 6,
                data = c(NA,NA,"log_age",NA, NA, NA,
                         NA, NA, NA, NA, NA, NA, 
                         "speed",NA,"accuracy",NA,"vocab",NA,
                         NA, NA, NA, NA, NA, NA, 
                         "log_rt","log_rt_sd", "acc",
                         "acc_sd", "prod","comp"), 
                byrow = TRUE)

node_text <- tibble(name=c("acc", "acc_sd", "accuracy", "comp", "log_age", "log_rt", "log_rt_sd", "prod", "speed", "vocab"), pretty=c("Accuracy", "Accuracy (SD)", "Accuracy", "Comprehension", "Age (log months)", "log RT", "log RT (SD)", "Production", "Speed", "Vocabulary"))

nodes <- get_nodes(fit3_age) |> tibble() |> left_join(node_text) |> mutate(label=ifelse(shape=="oval",pretty, str_c(pretty, "\n", est_sig)))

edges <- get_edges(fit3_age) |> filter(label!="1.00") |> mutate(label_location=.5) |> mutate(label_hjust=.5, label_vjust=.5, label_lineheight=1.5)

edges[9, "label_hjust"]=1
edges[7, "label_hjust"]=1

edges[18, "label_hjust"]=0
edges[16, "label_hjust"]=0

edges[17,"curvature"]=50
edges[17,"label_location"]=.6
edges[, "label_fill"]="white"

source(here("helper/modified_sem_plot.R"))

#View(edges)
prepare_graph(edges=edges, nodes=nodes, text_size = 2.5, layout = t(layout_age),
          rect_width=4.2,
          rect_height=3,
          ellipses_width=3,
          ellipses_height=3,
          variance_height=2,
          variance_width=1.5,
          arrow_angle=15,
          arrow_length=.1,
          var_arrow_angle=15,
          var_arrow_length=.1,
          spacing_y=4,
          spacing_x=4,
          fix_coord=FALSE) |> plot_happy_arrows()
```

Figure \ref{fig:age_regression} shows a regression model fit to this confirmatory factor analysis, with log age predicting each latent variable. This regression model allows interpretation of the covariances between latent factors as partial correlations (controlling for age). The non-age related variance of all three latent factors was significantly related to that of the other factors, with RT and accuracy showing strong negative covariance ($\beta=$ `r round(fit3_age_summary$pe[21,"est"],3)`, SE = `r round(fit3_age_summary$pe[21,"se"],3)`, $p < .0001$) and weaker but significant covariation between RT and vocabulary ($\beta=$ `r round(fit3_age_summary$pe[20,"est"],3)`, SE = `r round(fit3_age_summary$pe[20,"se"],3)`, $p < .0001$) and accuracy and vocabulary ($\beta=$ `r round(fit3_age_summary$pe[19,"est"],3)`, SE = `r round(fit3_age_summary$pe[19,"se"],3)`, $p < .0001$). This model supports the idea that variation in speed and accuracy of word recognition is related to differences in parent-reported vocabulary beyond the effects of age. These analyses all treat between-person differences, however, rather than modeling change in these factors within individuals.

## Speed of processing relates to vocabulary growth {.unnumbered}

```{r reliability}
MONTH_CUTOFF <- 3

longitudinal <- d_sub |>
  group_by(dataset_name, subject_id) |>
  count() |>
  filter(n > 1)

d_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(admin_num = 1:n(), 
         time_since_t0 = age - age[1],
         delta_t = c(0, diff(age)))

d_reliability <- d_long |>
  filter(time_since_t0 <= MONTH_CUTOFF, 
         admin_num <= 2) |>
  pivot_wider(id_cols = c("subject_id","dataset_name"), 
              names_from = "admin_num",
              values_from = c("log_rt", "log_rt_var", "long_window_accuracy", 
                              "long_window_acc_var", "prod", "comp"))

dataset_reliabilities <- d_reliability |>
  group_by(dataset_name) |>
  summarise(rt = cor(log_rt_1, log_rt_2, use = "pairwise.complete.obs"),
            acc = cor(long_window_accuracy_1, 
                          long_window_accuracy_2, use = "pairwise.complete.obs"))

# note this is not correct for CDIs because some CDIs can get grouped with two admins, inflating correlations
global_reliabilities <- d_reliability |>
  ungroup() |>
  summarise(rt = cor(log_rt_1, log_rt_2, use = "pairwise.complete.obs"),
            acc = cor(long_window_accuracy_1, 
                          long_window_accuracy_2, use = "pairwise.complete.obs"))
```

To probe within-person changes, we first examined test-retest reliability for our primary variables of interest by calculating Pearson correlations between pairs of administrations given no more than three months apart. Test-retest correlations were significant but relatively modest ($\rho_{acc} = `r round(global_reliabilities[2], 3)`$, $\rho_{rt} =  `r round(global_reliabilities[1],3)`$), suggesting that, across studies, LWL measures from individual sessions provide relatively noisy measurements of individual children even if individual studies are able to achieve higher reliability by averaging across measurement occasions [e.g., @marchman2008].

```{r}
d_growth <- d_sub |>
  filter(!is.na(prod), subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(rt_t0 = log_rt[1]) |>
  ungroup() |>
  mutate(rt_t0_group = ifelse(rt_t0 < median(rt_t0, na.rm=TRUE), "low", "high"))

d_growth$age_15 <- d_growth$age - 15

# note that it's singular to add an age by dataset random effect
growth_mod <- lmer(prod ~ poly(age_15,2) *rt_t0 + (age  | subject_id) + (1 | dataset_name), 
     data = d_growth, 
     control = lmerControl(optimizer = "bobyqa"))

# summary(growth_mod)
```

Given this relatively low reliability in our longitudinal datasets (which also tended to be from younger children), we chose to measure relationships between LWL and vocabulary using longitudinal growth models. We began by reproducing the analysis reported in @fernald2012, in which between-person differences in longitudinal growth in productive vocabulary were predicted based on between-person differences in RT during the initial session of the study. We fit a mixed effects model predicting growth in vocabulary as a quadratic function of age, RT at study initiation ($t_0$), and their interaction (as well as random effects of age by participant and by dataset). This model revealed a significant effect of $t_0$ RT (`r papaja::apa_print(growth_mod)$full_result$rt_t0`) and an interaction between $t_0$ RT and the quadratic age predictor (`r papaja::apa_print(growth_mod)$full_result$polyage_15_22_rt_t0`). This analysis suggests that children with faster initial RTs show both larger vocabularies and faster vocabulary growth over time, confirming that findings from @fernald2012 were robust across the full dataset. We confirmed this analysis using a non-linear growth model with a logistic shape, which shows a better fit to vocabulary size than the quadratic model (see SI: Non-Linear Growth Model). Figure \ref{fig:growth} shows predictions from this model, confirming the differentiation of growth curves for children with higher and lower initial reaction time.

```{r growth, fig.cap="Growth curves from a logistic growth model showing predicted vocabulary growth for the 25th and 75th percentile on initial reaction times. Solid lines show global model estimates and dashed lines show estimates for individual datasets with sufficient age variation as above.", fig.env="figure", fig.height = 2.5, fig.width=4, out.width=".5\\textwidth", include=TRUE}

d_sub_prod <- filter(d_sub, !is.na(prod)) |>
  group_by(subject_id) |>
  mutate(log_rt_0 = log_rt[1],
         acc_0 = long_window_accuracy[1]) |> 
  filter(!is.na(log_rt_0))

mod_nlme <- nlme(model = prod ~ SSlogis(age, Asym = 1, xmid, scale),
             data = d_sub_prod,
             fixed = xmid + scale ~ log_rt_0,
             random = xmid + scale ~ 1,
             groups = ~ dataset_name,
             start = c(xmid = 20, scale = 3, `xmid:log_rt_0` = 1, `scale:log_rt_0` = 3),
             na.action = na.exclude, 
             control = list(maxIter = 1000, tolerance = 10))

# fixed effects (no CIs)
global_preds <- expand_grid(age = seq(min(d_sub_prod$age), 
                                      max(d_sub_prod$age), .1), 
                            log_rt_0 = c(quantile(d_sub_prod$log_rt, 
                                                  c(.25,.75), 
                                                  na.rm=TRUE)))
global_preds$pred <- predict(mod_nlme, newdata = global_preds, level = 0)

#random effects
dataset_preds <- d_sub |>
  filter(!is.na(prod)) |> # filter on predictor
  group_by(dataset_name) |>
  summarise(min_age = min(age),
            max_age = max(age)) |>
  group_by(dataset_name) |>
  mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1), log_rt_0 = c(quantile(d_sub_prod$log_rt, 
                                                  c(.25,.75), 
                                                  na.rm=TRUE))))) |>
  select(-min_age, -max_age) |>
  unnest(col = "df") |>
  ungroup()

dataset_preds$pred <- predict(mod_nlme,
                              newdata = dataset_preds,
                              level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod, col = as.factor(log_rt_0))) + 
  geom_point(alpha = .02, color="black") + 
  geom_line(data = dataset_preds |>   
              filter(dataset_name %in% datasets_with_age_variation$dataset_name), 
            aes(y = pred, group = interaction(log_rt_0, dataset_name)), lty="dashed", alpha=.5)+
  geom_line(data = global_preds, aes(y = pred, group = log_rt_0, col = as.factor(log_rt_0)), size=1)+
  scale_color_viridis(discrete=T, option = "H")+
  theme(legend.position = "none")+
  labs(y="Production", x="Age in months")
# +
#   geom_line(data = dataset_preds, aes(y = pred))
```

```{r prep_long_data}
# rename and scale variables
d_sub_s <- d_sub |>
  ungroup() |>
  select(dataset_name, subject_id, administration_id, age, 
         time_since_t0, delta_t,
         log_rt, log_rt_var, 
         long_window_accuracy, long_window_acc_var, prod, comp, ) |>
  rename(acc = long_window_accuracy, 
         acc_sd = long_window_acc_var, 
         log_rt_sd = log_rt_var) |>
  ungroup() |>
  mutate(across(all_of(c("log_rt", "log_rt_sd", "acc", "acc_sd", 
                         "prod", "comp")), 
                       ~ age_scale(.x, age)))

d_sub_wide_fine <- d_sub_s |>
  mutate(t = cut(time_since_t0, breaks = c(0, 2, 4, 6, 8, 10, 30), 
                 include.lowest = TRUE)) |>
  filter(t != "(12,30]") |>
  mutate(t = as.numeric(t) -1) |>
  group_by(subject_id, dataset_name, t) |>
  summarise(across(all_of(c("log_rt", "log_rt_sd", "acc",
                            "acc_sd", "prod", "comp")),
                   ~ mean(.x, na.rm=TRUE))) |> 
  mutate(across(all_of(c("log_rt", "log_rt_sd", "acc",
                         "acc_sd", "prod", "comp")),
                ~ ifelse(is.nan(.x), NA, .x))) |>
  pivot_wider(id_cols = c("subject_id","dataset_name"), 
              names_from = "t",
              values_from = c("log_rt", "log_rt_sd", "acc", 
                              "acc_sd", "prod", "comp"), 
              names_prefix = "t")

slopes_model_long <- "

# regressions
accuracy_intercept =~ 1*acc_t0 + 1*acc_t1 + 1*acc_t2 + 1*acc_t3 + 1*acc_t4
accuracy_slope =~ 1*acc_t0 + 2*acc_t1 + 3*acc_t2 + 4*acc_t3 + 5*acc_t4
speed_intercept =~ 1*log_rt_t0 + 1*log_rt_t1 + 1*log_rt_t2 + 1*log_rt_t3 + 1*log_rt_t4
speed_slope =~ 1*log_rt_t0 + 2*log_rt_t1 + 3*log_rt_t2 + 4*log_rt_t3 + 5*log_rt_t4
vocab_intercept =~ 1*prod_t0 + 1*prod_t1 + 1*prod_t2 + 1*prod_t3 + 1*prod_t4
vocab_slope =~ 1*prod_t0 + 2*prod_t1 + 3*prod_t2 + 4*prod_t3 + 5*prod_t4
# let variances of each latent vary
accuracy_intercept ~~ NA*accuracy_intercept
accuracy_slope ~~ NA*accuracy_slope
speed_intercept ~~ NA*speed_intercept
speed_slope ~~ NA*speed_slope
vocab_intercept ~~ NA*vocab_intercept
vocab_slope ~~ NA*vocab_slope
"

slopes_long <- growth(slopes_model_long, d_sub_wide_fine, std.lv=TRUE, missing='fiml')

slopes_long_summary <- suppressMessages(summary(slopes_long, fit.measures=TRUE, standardize=TRUE))
```

On the other hand, it is possible that differences in predicted growth trajectories are due to overall coupling between vocabulary size and language processing, rather than a specifically predictive relationship between $t_0$ RT and vocabulary growth. To test this relationship, we used longitudinal structural equation models. We separated the longitudinal speed, accuracy, and vocabulary data into two-month bins spanning up to 10 months ($t_0, ..., t_4$) and fit individual growth across each of these variables. We used full-information maximum likelihood to handle the substantial missing data caused by the different longitudinal sampling schemes of studies in our dataset. (We present a model of coupled growth in the observed variables here for simplicity; see SI for a comparable model using growth in the latent factors). The fitted longitudinal model is shown in Figure \ref{fig:longitudinal}. Overall fit statistics were generally acceptable (Confirmatory fit index: `r round(slopes_long_summary$fit["cfi"], 3)`, RMSE: `r round(slopes_long_summary$fit["rmsea"], 3)`, RMSE $p$-value: `r round(slopes_long_summary$fit["rmsea.pvalue"], 3)`).

Our key question of interest concerned coupling among the (latent) intercepts and slopes of these growth models. Consistent with the idea that overall faster processing is related to vocabulary growth, we saw significant coupling between processing speed intercepts and vocabulary growth slopes ($\beta=$ `r round(slopes_long_summary$pe[63,"est"],3)`, SE = `r round(slopes_long_summary$pe[63,"se"],3)`, $p =$ `r round(slopes_long_summary$pe[63,"pvalue"],3)`) as well as a variety of other couplings. On the other hand, there was not significant coupling between *growth* in RT and *growth* in vocabulary ($\beta=$ `r round(slopes_long_summary$pe[65,"est"],3)`, SE = `r round(slopes_long_summary$pe[65,"se"],3)`, $p =$ `r round(slopes_long_summary$pe[65,"pvalue"],3)`). This null effect could be interpreted as being consistent with these abilities growing independently, but we cannot rule out other possibilities. First, the longitudinal data we had might not allow sufficiently precise estimates of growth slopes, or second, since vocabulary growth is non-linear, the linear model we used here might not have captured coupling among nonlinear aspects of developmental change.

In sum, these findings provide evidence consistent with the claim that differences in processing speed are related to differences in the rate of age-related change in vocabulary [@fernald2006;@Weisleder2013]. Children with greater skill in word recognition learn words faster.

```{r longitudinal, fig.env="figure*", height = 3.5, fig.cap = "Structural equation model showing longitudinal couplings between growth parameters. "}
layout <- read.csv(here("misc/layout_slopes_observed_growth.csv"), header = FALSE)

acc_color <- "#B2DF8A"
rt_color <- "#FB9A99"
prod_color <- "#FFFF99"

intercept_color <- "#1F78B4"
intercept_background <- "#A6CEE3"
slope_color <- "#FF7F00"
slope_background <- "#FDBF6F"

#"#A6CEE3" "#1F78B4" "#B2DF8A" "#33A02C"
# [5] "#FB9A99" "#E31A1C" "#FDBF6F" "#FF7F00"
# [9] "#CAB2D6" "#6A3D9A" "#FFFF99" "#B15928"
node_text <- tribble(~name, ~ pretty, ~fill, 
                     "acc_t0", "t0 Accuracy", acc_color,
                     "acc_t1", "t1 Accuracy", acc_color,
                     "acc_t2", "t2 Accuracy",acc_color,
                     "acc_t3", "t3 Accuracy",acc_color,
                     "acc_t4", "t4 Accuracy",acc_color,
                     "accuracy_intercept", "Accuracy intercept",acc_color,
                     "accuracy_slope", "Accuracy slope",acc_color,
                     "log_rt_t0", "t0 log RT",rt_color,
                     "log_rt_t1", "t1 log RT",rt_color,
                     "log_rt_t2", "t2 log RT",rt_color,
                     "log_rt_t3", "t3 log RT",rt_color,
                     "log_rt_t4", "t4 log RT",rt_color,
                     "prod_t0" , "t0 Production", prod_color,
                     "prod_t1" , "t1 Production",  prod_color,
                     "prod_t2" , "t2 Production" ,  prod_color,
                     "prod_t3", "t3 Production",         prod_color,
                     "prod_t4", "t4 Production" , prod_color,
                     "speed_intercept" , "Speed intercept" , rt_color,
                     "speed_slope" , "Speed slope",rt_color,
                     "vocab_intercept", "Vocabulary intercept", prod_color,
                     "vocab_slope", "Vocabulary slope",  prod_color ) |> 
  mutate(alpha=.1)

nodes <- get_nodes(slopes_long) |> tibble() |> left_join(node_text) |> mutate(label=pretty)

edges <- get_edges(slopes_long) |> mutate(label_location=.5) |> mutate(label=ifelse(arrow=="last", NA, label)) |>  mutate(connect_from=ifelse(to==from, "right",NA),
                    connect_to=connect_from) |> 
  mutate(color="black", label_hjust=.5, label_location=.5, label_fill="white")

edges[52,"curvature"]=85
#edges[52,"color"]="blue"
edges[61,"curvature"]=85
#edges[61,"color"]="blue"
edges[66,"curvature"]=85
#edges[66,"color"]="blue"
edges[52, "label_hjust"]=0
edges[61, "label_hjust"]=0
edges[66, "label_hjust"]=0

edges[54,"curvature"]=85
edges[54, "label_hjust"]=0

edges[63,"curvature"]=85
edges[63, "label_hjust"]=0

edges[53,"curvature"]=80
edges[53,"color"]=intercept_color
edges[53, "label_fill"]=intercept_background
edges[55,"curvature"]=82
edges[55,"label_location"]=.55

edges[55,"color"]=intercept_color
edges[55, "label_fill"]=intercept_background

edges[62,"curvature"]=80
edges[62,"color"]=intercept_color
edges[62, "label_fill"]=intercept_background

edges[53, "label_location"]=.6
edges[53, "label_hjust"]=0.4

edges[58,"curvature"]=80
edges[58,"color"]=slope_color
edges[58, "label_fill"]=slope_background

edges[60,"curvature"]=82
edges[60,"color"]=slope_color
edges[60, "label_fill"]=slope_background
edges[60,"label_location"]=.45
edges[65,"curvature"]=80
edges[65,"color"]=slope_color
edges[65, "label_fill"]=slope_background
edges[65, "label_location"]=.4
edges[53, "label_hjust"]=0.4

edges[59,"curvature"]=77

edges[64,"curvature"]=75
edges[56,"curvature"]=75
edges[57,"curvature"]=75

source(here("helper/modified_sem_plot.R"))
#View(edges)
prepare_graph(edges=edges, nodes=nodes, text_size = 2, layout = t(layout),angle=0,
          rect_width=1.5,
          rect_height=2,
          ellipses_width=1.5,
          ellipses_height=3,
          variance_width= .5,
          variance_height=1.3,
          var_arrow_angle=10,
          var_arrow_length=.05,
          arrow_angle=15,
          arrow_length=.07,
          variance_diameter=1,
          spacing_y=2.5,
          spacing_x=1.5,
          fix_coord=FALSE) |> 
  edit_edges({connect_from=case_when(
    (to==from & from %in% c("accuracy_intercept","accuracy_slope","speed_slope",
                          "speed_intercept", "vocab_slope", "vocab_intercept"))~"bottom",
    to==from ~ "right",
    !is.na(curvature)~"left",
    T ~ "right")}) |> 
  plot_happy_arrows()
```

# Discussion {.unnumbered}

How does word recognition change across early childhood and how does it relate to language learning? We investigated these questions using a new, large-scale dataset of developmental eye-tracking measurements compiled across many prior studies. The age-gradients for speed and accuracy indicate that both improved asymptotically. Gradients for recognition speed were consistent with the log-log relationship associated with the "power law of practice", that is, with a gradual convergence on mature levels of processing efficiency. Further, the age gradient suggested that trial-to-trial variability decreases with age, consistent with both the literature on skill learning [@todorov2002optimal] and other work on developmental changes in variability [@judd2021working;@galeano2018variability;@kautto2023introducing]. Speed and accuracy were both related to vocabulary size concurrently and processing speed was also related to later vocabulary growth.

Together, our findings are consistent with theories that posit that language learning is a process of skill acquisition, in which children become adept at quickly converting ephemeral signals into meaning [@christiansen2016]. This skill develops gradually over the course of early childhood and supports word learning. Further, our results point to consistency between skill development in early childhood and the continued refinement of language processing and language knowledge during middle childhood [@mcmurray2022slow;@colby2023efficiency].

By aggregating data from many pre-existing studies, we were able to overcome the limitations of prior investigations, which typically had sample sizes at least an order of magnitude smaller than ours. In contrast to individual studies, which typically have at best the statistical power to test one or two specific contrasts, our "big data" approach provided the sample sizes necessary to explore the relationships between different variables. Because early language is so variable [@frank2021], these kinds of samples -- with thousands, rather than dozens of children -- are likely to be required to gain further insight into the psychometrics of early language learning [@bergelson2023;@frank2017;@frank2021;@manybabies2020]. Future work must also explore how the functional forms we observed here between individuals reflect processes of within-person change. Although the Peekbank dataset includes a variety of longitudinal data, most reflect a small number of measurements; denser longitudinal data collection is required to better estimate  within-person growth models.

At the same time, our approach is both observational and exploratory. Thus, we cannot untangle the range of different causal models that explain the variation we observed. First, early word recognition skill could lead to faster word learning, due to skilled children gleaning more information from the same signal [@fernald2006]. But second, faster children could also be faster due to their larger vocabulary and stronger lexical representations. These two causal directions could also interact reciprocally, leading to a "rich get richer" process in which children with larger vocabularies process faster, and their faster processing helps them increase their vocabulary size more rapidly. Finally, a third shared factor -- perhaps general cognitive ability -- could underpin both processes. Our cross-sectional data cannot distinguish these hypotheses even in principle [@vanderweele2023dimensional], and our longitudinal data are likely too sparse to distinguish such complex causal models. Thus, these questions remain an open target for dense longitudinal data collection.

Our findings here are limited in their generalizability by the convenience samples that were used in the individual studies in the Peekbank database. These studies typically (but not always) represent children from well-educated parents living in university-adjacent communities. We would not expect that specific numerical parameters estimated in this aggregate convenience sample would generalize to other samples. Nevertheless, the consistency of the trends we observed across datasets suggests that our qualitative conclusions are robust to some significant sociodemographic variation.

More broadly, our results here suggest the continued importance of the looking-while-listening paradigm as an index of children's language processing abilities. If language learning is, at least in part, a process of skill learning, then measurement of this skill is a critical window into understanding the remarkable process of language learning.

<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak

\newpage