### 2025-04-14 paper/supplemental_information.Rmd
prep for submission

---
title: "Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes:
  - \usepackage{threeparttablex}
  - \usepackage{booktabs}
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
library(here)
source(here("helper","common.R"))
```

```{r, echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE}
paper <- suppressMessages(knitr::purl(here("paper","paper.Rmd"), documentation = 0))
suppressMessages(source(paper))
```

# Dataset Description

Figure \ref{fig:dataset} gives the age distribution of unique participants for each separate dataset at different ages. 
Note that for some datasets, there are multiple administrations (i.e., experimental test sessions) for each participant.

```{r dataset, fig.cap="\\label{fig:dataset} Age distribution for each dataset, using three-month bins.", fig.pos="h!"}

# summarize participants by age bin
d_sub_admin_by_age <- d_sub |>
  mutate(age_bin = age %/% 3 * 3+1.5) |>
  group_by(dataset_name, age_bin) |>
  summarise(
    n_subj = n_distinct(subject_id),
    n_administration = n_distinct(administration_id)) |>
  ungroup()

# ggplot(d_sub, aes(x = age))+ 
#   geom_histogram(binwidth=3) + 
#   facet_wrap(~dataset_name, scale = "free_y") +
#   xlab("Age (months)") + 
#   ylab("Number of administrations")+
#   theme(strip.text.x = element_text(size = 5))

# plot number of participants across age
ggplot(d_sub_admin_by_age, aes(x = age_bin,y=n_subj))+ 
  geom_bar(stat="identity",width=3) + 
  facet_wrap(~dataset_name, scale = "free_y") +
  xlab("Age (months)") + 
  ylab("Number of participants")+
  theme(strip.text.x = element_text(size = 5))

```

Figure \ref{fig:longitudinal_descriptives} shows the distribution of measurement intervals for longitudinal studies within the dataset.

```{r longitudinal_descriptives, fig.cap="\\label{fig:longitudinal_descriptives} Number of administrations for datasets with repeated measurements, plotted by dataset.", fig.pos="h!"}

d_sub_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(admin_num = 1:n(), 
         delta_t = age - age[1])

ggplot(d_sub_long, 
       aes(x = delta_t, fill = dataset_name)) +
  geom_histogram(binwidth = 1) + 
  scale_fill_solarized(name = "Dataset") +
  theme(legend.position = "bottom") + 
  # facet_wrap(~dataset_name) + 
  xlab("Age since first test") + 
  ylab("Number of participants")
```

# Test-Retest Reliability

```{r reliability}
MONTH_CUTOFF <- 3

longitudinal <- d_sub |>
  group_by(dataset_name, subject_id) |>
  count() |>
  filter(n > 1)

d_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(admin_num = 1:n(), 
         time_since_t0 = age - age[1],
         delta_t = c(0, diff(age)))

d_reliability <- d_long |>
  filter(time_since_t0 <= MONTH_CUTOFF, 
         admin_num <= 2) |>
  pivot_wider(id_cols = c("subject_id","dataset_name"), 
              names_from = "admin_num",
              values_from = c("log_rt", "long_window_accuracy", 
                              "short_window_accuracy"))

dataset_reliabilities <- d_reliability |>
  group_by(dataset_name) |>
  summarise(rt = cor(log_rt_1, log_rt_2, use = "pairwise.complete.obs"),
            acc_long = cor(long_window_accuracy_1, 
                           long_window_accuracy_2, use = "pairwise.complete.obs"),
            acc_short = cor(short_window_accuracy_1, 
                            short_window_accuracy_2, use = "pairwise.complete.obs"))

global_reliabilities <- d_reliability |>
  ungroup() |>
  summarise(rt = cor(log_rt_1, log_rt_2, use = "pairwise.complete.obs"),
            acc_long = cor(long_window_accuracy_1, 
                           long_window_accuracy_2, use = "pairwise.complete.obs"),
            acc_short = cor(short_window_accuracy_1, 
                            short_window_accuracy_2, use = "pairwise.complete.obs"))
```

We examined test-retest reliability for our primary variables of interest by calculating Pearson correlations between pairs of administrations given no more than three months apart. Test-retest correlations were significant but relatively modest: $\rho_{long window acc} = `r round(global_reliabilities[2], 3)`$, $\rho_{short window acc} = `r round(global_reliabilities[3], 3)`$, $\rho_{rt} =  `r round(global_reliabilities[1],3)`$. These reliabilities were biased downwards by three factors, however. First, longitudinal assessments sometimes use variable items between testing sessions, leading to item-related variance in measurement. Second, even three months can lead to substantial change in some children's language abilities, thus correlations are attenuated by true change as well as measurement error. Third, longitudinal data in the dataset come primarily from the youngest children and hence are likely to show overall higher measurement error due to variability in children's behavior and an overall lower number of trials.

# Correlations

Table \ref{tab:corrs} shows pairwise correlations between the primary variables of interest in the dataset.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{llllllllll}
\toprule
& \multicolumn{1}{l}{age} & \multicolumn{1}{l}{log age} & \multicolumn{1}{l}{RT} & \multicolumn{1}{l}{log RT} & \multicolumn{1}{l}{long acc} & \multicolumn{1}{l}{short acc } & \multicolumn{1}{l}{prod} & \multicolumn{1}{l}{comp}\\
\midrule
age &  1.00 &  &  &  &  &  &  & \\
log age  & 0.98 & 1.00 &  &  &  &  &  &\\
rt  & -0.33 & -0.35 & 1.00 &  &  &  &  & \\
log rt  & -0.34 & -0.36 & 0.96 & 1.00 &  &  &  & \\
long window accuracy  & 0.44 & 0.48 & -0.48 & -0.46 & 1.00 &  &  & \\
short window accuracy & 0.38 & 0.43 & -0.62 & -0.61 & 0.82 & 1.00 &  & \\
production vocabulary & 0.72 & 0.70 & -0.31 & -0.33 & 0.51 & 0.45 & 1.00 & \\
comprehension vocabulary & 0.42 & 0.42 & -0.25 & -0.24 & 0.24 & 0.24 & 0.59 & 1.00\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Pairwise correlations between primary variables of interest. \label{tab:corrs}}
\end{table}

```{r corrs, results = "asis", include = FALSE}

cor(select(d_sub, age, log_age, rt, log_rt, 
           long_window_accuracy, short_window_accuracy, prod, comp), 
    use = "pairwise.complete.obs") |>
  round(2) |>
  papaja::apa_table()

```

# Functional Form Model Comparison

Table \ref{tab:accmods} shows model comparison measures between for different models of the functional form of the relationship between accuracy and age and Table \ref{tab:rtmods} shows the same for reaction time. Age gradients are estimated substantially better with long window accuracies. Note that there are a greater number of observations for short window accuracies due to less missing data. We speculate that, on average, more participants looked away from the screen towards the end of trials, leading to a greater number of exclusions of long window trials based on the 50% criterion. Note that the total percentage of trials excluded is still small for both measures: `r round(mean(is.na(d_trial$long_window_accuracy)), 2)`% for long window accuracy and `r round(mean(is.na(d_trial$short_window_accuracy)), 2)`% for short window accuracy.

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{lllllllll}
\toprule
n. obs & \multicolumn{1}{c}{sigma} & \multicolumn{1}{c}{logLik} & \multicolumn{1}{c}{AIC} & \multicolumn{1}{c}{BIC} & \multicolumn{1}{c}{REMLcrit} & \multicolumn{1}{c}{df.residual} & \multicolumn{1}{c}{model} & \multicolumn{1}{c}{$r_2$}\\
\midrule
48354 & 0.28 & -8,338.83 & 16,691.65 & 16,753.16 & 16,677.65 & 48347 & Long window, linear age & 0.14\\
48354 & 0.28 & -8,305.39 & 16,624.78 & 16,686.28 & 16,610.78 & 48347 & Long window, log age & 0.12\\
50244 & 0.31 & -12,486.60 & 24,987.21 & 25,048.98 & 24,973.21 & 50237 & Short window, linear age & 0.09\\
50244 & 0.31 & -12,458.77 & 24,931.54 & 24,993.31 & 24,917.54 & 50237 & Short window, log age & 0.08\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Model comparison metrics for different functional forms of the relationship between accuracy and age. \label{tab:accmods}}
\end{table}

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{lllllllll}
\toprule
n. obs & \multicolumn{1}{c}{sigma} & \multicolumn{1}{c}{logLik} & \multicolumn{1}{c}{AIC} & \multicolumn{1}{c}{BIC} & \multicolumn{1}{c}{REMLcrit} & \multicolumn{1}{c}{df.residual} & \multicolumn{1}{c}{model} & \multicolumn{1}{c}{$r_2$}\\
\midrule
18940 & 0.45 & -12,414.69 & 24,843.38 & 24,898.32 & 24,829.38 & 18933 & Log RT, linear age & 0.21\\
18940 & 0.45 & -12,393.27 & 24,800.53 & 24,855.48 & 24,786.53 & 18933 & Log RT, log age & 0.21\\
18940 & 570.49 & -147,644.43 & 295,302.87 & 295,357.81 & 295,288.87 & 18933 & Linear RT, linear age & 0.18\\
18940 & 570.02 & -147,622.67 & 295,259.34 & 295,314.29 & 295,245.34 & 18933 & Linear RT, log age & 0.18\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Model comparison metrics for different functional forms of the relationship between RT and age. \label{tab:rtmods}}
\end{table}

Note that with our current dataset we cannot specifically determine whether within-individual patterns of change conform to linear, power law, or exponential developmental patterns (Heathcote, Brown, \& Mewhort, 2000). Our current results apply to the form of the age gradient as opposed to the form of any individual's pattern of developmental change.

```{r accmods, results="asis", include = FALSE}
papaja::apa_table(acc_mods_lmer_summary)
```

```{r results="asis", include = FALSE}
papaja::apa_table(rt_mods_lmer_summary)
```

# Factor Analysis

Figure 3 shows the result of a parallel analysis supporting the presence of three factors in the exploratory factor analysis. Table \ref{tab:loadings} shows the factor loadings for the exploratory three-factor solution using varimax rotation. The first factor is primarily driven by vocabulary measures, the second by reaction time, and the third by accuracy measures.

```{r parallel, fig.cap="Parallel analysis scree plot showing the eigenvalues for each factor, for actual, simulated, and resampled data."}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(dataset_name, rt, rt_var, long_window_accuracy, long_window_acc_var, prod, comp, age)

d_sub_mat_s <- d_sub_mat |>
  ungroup() |>
  mutate(across(all_of(c("rt", "rt_var", "long_window_accuracy", 
                         "long_window_acc_var", "prod", "comp")), 
                ~ age_scale(.x, age))) 
outputs <- capture.output(fa.parallel(select(d_sub_mat, -dataset_name, -age), fa = "fa", 
            use = "pairwise.complete.obs"))
```

```{r eval=FALSE}
fs <- fa(select(d_sub_mat, -dataset_name), nfactor = 3,
         use = "pairwise", rotate = "varimax")

as.data.frame(unclass(loadings(fs))) |> 
  xtable(caption = "Factor loadings for the exploratory three factor solution using varimax rotation.", digits = 2)
```

<!-- % latex table generated in R 4.4.2 by xtable 1.8-4 package -->
<!-- % Mon Apr 14 12:25:14 2025 -->
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & F1 & F2 & F3 \\ 
  \hline
RT & -0.19 & 0.81 & -0.30 \\ 
  RT var & -0.10 & 0.81 & -0.22 \\ 
  long window accuracy & 0.33 & -0.31 & 0.55 \\ 
  long window accuracy var & -0.10 & 0.26 & -0.65 \\ 
  production vocabulary & 0.95 & -0.04 & 0.30 \\ 
  comprehension vocabulary & 0.61 & -0.16 & -0.01 \\ 
  age & 0.63 & -0.14 & 0.37 \\ 
   \hline
\end{tabular}
\caption{\label{tab:loadings}Factor loadings for the exploratory three factor solution using varimax rotation.} 
\end{table}

# Alternative Factor Structures

In this section, we provide comparisons between the three-factor model we report in the main text and several alternative models, including:

* a one-factor model;
* a two-factor model with vocabulary separated from speed and accuracy;
* a two-factor model with speed separated from accuracy and vocabulary; and 
* a two-factor model with variability terms separated from speed, accuracy, and vocabulary.

\noindent Table \ref{tab:modcomp} shows the result of these comparisons. The three-factor model shows the lowest AIC and BIC, as well as being significantly better fittign than the next-best model.

```{r, eval=FALSE}
fa1_model <-  "
F1  =~ log_rt + log_rt_var + long_window_accuracy + long_window_acc_var + prod + comp"
fa1_fit <- cfa(fa1_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

fa2_model <-  "
processing =~ log_rt + log_rt_var + long_window_accuracy + long_window_acc_var
vocabulary =~ prod + comp"
fa2_fit <- cfa(fa2_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

fa2b_model <-  "
accuracy =~ long_window_accuracy + long_window_acc_var + prod + comp
speed =~ log_rt + log_rt_var"
fa2b_fit <- cfa(fa2b_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

fa2v_model <-  "
language =~ log_rt + long_window_accuracy + prod + comp
               variability =~  long_window_acc_var + log_rt_var"
fa2v_fit <- cfa(fa2v_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

fa3_model <-  "
speed =~ log_rt  + log_rt_var
accuracy =~ long_window_accuracy + long_window_acc_var 
vocab =~ prod + comp"
fa3_fit <- cfa(fa3_model, d_sub_mat_s, std.lv=TRUE, missing='fiml')

xtable(anova(fa1_fit, fa2_fit, fa2b_fit, fa2v_fit, fa3_fit), digits = 2)
```
<!-- xtable 1.8-4 package -->
<!-- % Fri Apr  4 15:15:52 2025 -->
\begin{table}[ht]
\centering
\caption{Model comparison for alternative factor structures. $p$-values show differences between adjacent models; no $p$-values are shown for comparisons between non-nested models. \label{tab:modcomp}}
\begin{tabular}{lrrrrrrrr}
  \hline
 & Df & AIC & BIC & Chisq & Chisq diff & RMSEA & Df diff & Pr($>$Chisq) \\ 
  \hline
Three-factor  & 6 & 46346.69 & 46475.43 & 91.96 &  &  &  &  \\ 
  Two-factor (vocab)  & 8 & 46397.74 & 46514.22 & 147.01 & 55.05 & 0.09 & 2 & $<$ 0.0001 \\ 
  Two-factor (speed) & 8 & 46486.96 & 46603.44 & 236.23 & 89.22 & 0.00 & 0 &  \\ 
  Two-factor (variability) & 8 & 46536.10 & 46652.58 & 285.37 & 49.14 & 0 & 0 &  \\ 
  One-factor & 9 & 46535.49 & 46645.84 & 286.76 & 1.39 & 0.01 & 1 & 0.24 \\ 
   \hline
\end{tabular}
\end{table}

# Non-linear Growth Models

To test for the differentiation of vocabulary growth based on initial reaction time, we used the package `nlme` to fit a logistic growth model to the production data. This model has two parameters for the logistic curve, a scale and an intercept. Both were allowed to interact with initial reaction time. We also included random effects of logistic intercept and scale by participant and a grouping term across datasets. This model showed a significant effect of initial reaction time on the intercept of the logistic growth curve, but not on its scale (see Table \ref{tab:nlme}). 
```{r}
# xtable(summary(mod_nlme)$tTable)
```

<!-- % latex table generated in R 4.4.2 by xtable 1.8-4 package -->
<!-- % Thu Mar 13 13:10:30 2025 -->
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
& Value & Std Error & DF & $t$-value & $p$-value \\ 
\hline
Growth Intercept & 2.26 & 2.70 & 1766.00 & 0.84 & 0.40 \\ 
Growth Intercept $\times$ $t_0$ RT & 3.30 & 0.38 & 1766.00 & 8.59 & < 0.0001 \\ 
Growth Scale & 2.05 & 2.60 & 1766.00 & 0.79 & 0.43 \\ 
Growth Scale $\times$ $t_0$ RT & 0.28 & 0.38 & 1766.00 & 0.75 & 0.46 \\ 
\hline
\end{tabular}
\caption{Fixed effects estimates from logistic growth model. \label{tab:nlme}}
\end{table}

Interpretation of growth in both this model and the linear growth model in the main text is complicated by the fact that the CDI form puts a ceiling on the total number of words that can be recorded; both the quadratic growth functions and the logistic functions come together at the form ceiling. Thus, a shift in quadratic growth in the linear model and a shift in intercept in the logistic model both point to the same overall effect, which is faster growth at the point of maximal sensitivity of the CDI. Neither model can estimate whether the overall growth trajectory is different beyond the range of the CDI. Thus, although these models might initially seem to be in conflict, we believe that they actually point to the same phenomenon, which is perhaps better described by the longitudinal SEM model reported in the main text. Children with greater skill in word recognition show an overall positive shift in the growth trajectory of vocabulary development.

# References

Heathcote, A., Brown, S., \& Mewhort, D. J. K. (2000). The power law repealed: The case for an exponential law of practice. Psychonomic Bulletin & Review, 7(2), 185–207.